---
title: "pseudobulk_CMP"
output: html_notebook
---

# Purpose
Determine _ranked spearman correlation_ between single cell RNASeq (scRNASeq) and bulk RNASeq (bRNASeq)

## Method
1. Perform pseudo-bulk analysis from scRNASeq of population (i.e., CMP)
2. Download bRNASeq of same population (has 2 replicates)
3. Ranked spearman correlation between bRNASeq replicates
4. Ranked spearman correlation between each bRNASeq replicate and pseudobulk scRNASeq sample
5. Ranked spearman correlation between combined bRNASeq replicates and pseudobulk scRNASeq sample


# Notebook setup

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

## Load libraries
```{r setup}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(SingleCellExperiment)
library(Matrix.utils)
library(DESeq2)
library(edgeR)
library(csaw)
library(tximport)
```

## Set global variables
```{r}
projectName <- "pseudobulk_CMP"
```

## Store session info
```{r}
sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
sink(sessionInfo.filename)
sessionInfo()
sink()
```

## Load local scripts
```{r}
# source("../RFunctions/read_10XGenomics_data.R")
# source("../RFunctions/PercentVariance.R")
source("../RFunctions/Mouse2Human_idconversion.R")
source ("../RFunctions/ColorPalette.R")
```

# Load scRNASeq object

## Dim 25
Using CMP.rds object of dim=25, describing 85% of variation
```{r}
# CMP seurat object generated in "CMPSubset.Rmd"
seurat.object <- readRDS("CMP_dim25.RDS")
```
```{r fig.height=5, fig.width=15}
p <- FeaturePlot(seurat.object, features = c("Fli1", "Klf1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
plot(p)
png(filename = "CMP_Fli1Klf1Ratio.png", width = 2400, height = 400)
plot(p)
dev.off()
```

```{r fig.height=5, fig.width=15}
p <- FeaturePlot(seurat.object, features = c("Gata2", "Gata1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
plot(p)
png(filename = "CMP_Gata2Gata1Ratio.png", width = 2400, height = 400)
plot(p)
dev.off()
```


## Calculate pseudobulk gene expression

__ Based on tutorial @ https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html __
```{r}
# Extract raw counts and metadata to create SingleCellExperiment object

counts <- seurat.object@assays$RNA@counts 

metadata <- seurat.object@meta.data

# Set up metadata as desired for aggregation and DE analysis
metadata$cluster_id <- factor(seurat.object@active.ident)

# Create single cell experiment object
sce <- SingleCellExperiment(assays = list(counts = counts), 
                           colData = metadata)

```

Investigate the object
```{r}
print(dim(counts(sce)))
print(dim(colData(sce)))
print(head(colData(sce)))
```

Note that QC was performed during Seurat analysis

```{r}
# sce <- perCellQCMetrics(sce)
```
## Aggregate counts
### Aggregate top-level counts

```{r}
groups <- colData(sce)[, "orig.ident"]
aggr.counts <- aggregate.Matrix(t(counts(sce)), groupings = groups, fun = "sum")

dim(aggr.counts)
```


```{r}
sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)

```





# Load bRNASeq data


```{r}
# import rsem.genes.results files

setwd("../../bRNASeq/")
file_list <- list.files(pattern = "^CMP", recursive = TRUE, include.dirs = TRUE)
names(file_list) <- lapply(file_list, function(x) gsub('.genes.results', '', x))
file_list
```

```{r}
mstxi <- tximport(files = paste0("../../bRNASeq/", file_list), type = 'rsem', txIn = FALSE, txOut = FALSE, geneIdCol = 3, )
colnames(mstxi$counts) <- gsub("^.*/", "", names(file_list))
```

```{r}
rna.counts <- merge(mstxi$counts, t(aggr.counts), by = "row.names", all = TRUE)
rna.counts[is.na(rna.counts)]<-0
rownames(rna.counts) <- rna.counts$Row.names
rna.counts <- rna.counts[, 2:ncol(rna.counts)]
head(rna.counts)
```

```{r}
dgelist_groups <- factor(c(unlist(stringr::str_extract_all(gsub("/.*", "", names(file_list)), pattern = "[a-zA-Z]+")), "SingleCell"))
levels(dgelist_groups)
```

```{r}
dgelist <- DGEList(counts = rna.counts, group = dgelist_groups)
dgelist
apply(dgelist$counts, 2, sum)
keep <- rowSums(cpm(dgelist) > 10) >=2
dgelist <- dgelist[keep,]
```


### Calculate library size
```{r}
dgelist$samples$lib.size <- colSums(dgelist$counts)
dgelist <- calcNormFactors(dgelist)
```

### Calculate dispersion

designMat
```{r}
designMat <- model.matrix(~ 0 + dgelist$samples$group)
colnames(designMat) <- levels(dgelist$samples$group)
designMat
```

estimate GLM dispersion and apply
```{r}
dgelist <- estimateGLMCommonDisp(dgelist, designMat)
dgelist <- estimateGLMTrendedDisp(dgelist, designMat, method = 'bin.spline')
dgelist <- estimateGLMTagwiseDisp(dgelist, designMat)
```

```{r}
fit <- glmQLFit(dgelist, designMat, robust = TRUE)
plotQLDisp(fit)
```

```{r}
qlf <- glmQLFTest(fit, coef = )
summary(decideTests(qlf))
```


#### rld on  individual samples

```{r}
rld <- rlog(round(dgelist$counts), blind = TRUE)
rld <- as.data.frame(rld)
rld.spear <- cor(rld, method = "spearman")
rld.melt <- reshape2::melt(rld)
colnames(rld.melt) <- c("sample", "rld")
```

```{r}
rld.spear
```



```{r}
ggplot(data = rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
```


```{r}
ggplot(data = rld, aes(x = CMPm2, y = CMP1010)) + geom_point()
```


#### rld on pooled samples

```{r}
pool.counts <- dgelist$counts
pool.counts <- as.data.frame(pool.counts)
pool.counts$ScriptSeq <- (pool.counts$CMP1009 + pool.counts$CMP1010)/2
pool.counts$TruSeq <- (pool.counts$CMP443 + pool.counts$CMP448)/2
pool.counts <- subset(pool.counts, select = c(ScriptSeq, TruSeq, CMPm2))
pool.counts <- as.matrix(pool.counts)
```


```{r}
pool.rld <- rlog(round((pool.counts)), blind = TRUE)
pool.rld.spear <- cor(pool.rld, method = "spearman")
pool.rld.melt <- reshape2::melt(pool.rld)
colnames(pool.rld.melt) <- c("sample", "rld")
```

```{r}
pool.rld.spear
```



```{r}
ggplot(data = pool.rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
```


```{r}
ggplot(data = pool.rld.melt, aes(x = CMPm2, y = ScriptSeq)) + geom_point()
```



Combine replicates from CPM table	

```{r}
rna.cpm <- as.data.frame(cpm(dgelist, log = TRUE, normalized.lib.sizes = TRUE))
rna.cpm$ScriptSeq <- (rna.cpm$CMP1009 + rna.cpm$CMP1010)/2
rna.cpm$TruSeq <- (rna.cpm$CMP443 + rna.cpm$CMP448)/2
rna.cpm <- subset(rna.cpm, select = c(ScriptSeq, TruSeq, CMPm2))
```

plot correlation and scatter grams
```{r}
cpm.spear <- cor(rna.cpm, method = "spearman")
cpm.melt <- reshape2::melt(rld)
colnames(cpm.melt) <- c("sample", "rld")
```

```{r}
cpm.spear
```








## Create DESeq object from single cell data
```{r}
sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)
```




